Distributional Trees and Forests

### 2017-11-24 pkg/disttree/inst/draft/draft.Rnw
small corrections
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1757 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 80fa692cfa1a87e6b2d6eb4ca6acabd3ffcd6f78

\Abstract{
In the field of probabilistic forecasting linear models such as ensemble model output statistics (EMOS, \citealt{gneiting2005calibrated}) provide common and widely used tools. This is due to their simplicity and their interpretability. However, the restriction to model only linear effects requires further extensions which allow for more flexibility. Generalized additive models for location, scale, and shape (GAMLSS, \citealt{stasinopoulos2005}) can also capture non-linear effects as each distribution parameter can be modeled separately by its own generalized additive model (GAM). A different approach that would make it possible to also model non-additive effects and interactions is to apply tree structured models such as regression trees (\citealt{loh2011classification}) or random forests (\citealt{breiman2001random}). However, up to now, this class of models has not yet been integrated in probabilistic forecasting as no general distributional framework is available for these methods. Therefore, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. Applying these novel approaches on precipitation data in a mountainous region with a large number of numerical weather prediction quantities illustrates their strengths as their performance is compared to GAMLSS models that are specified based either on subject-matter knowledge/experience from meteorology or based on a computationally more demanding boosting approach.

@@

\SweaveOpts{eval = TRUE}
\section{Introduction}
A typical application in probabilistic forecasting in meteorology is the analysis of precipitation data. The two main interests lie in estimating the probability of precipitation as well as predicting the expected amount of precipitation. In this research area various methods have been developed and tested in the last decades. (Multiple) Linear models such as ensemble model output statistics (EMOS) were introduced more than a decade ago (\citealt{gneiting2005calibrated}) and are still common and widely used. Despite their simplicity and interpretability, the limitation to modeling only linear effects has risen the interest in applying more flexible models. This feature is provided by generalized additive models (GAM) which can also model non-linear effects. However, often only the expectation of the underlying distribution is estimated. To obtain probabilistic predictions, generalized additive models for location, scale, and shape (GAMLSS, \citealt{stasinopoulos2005}) offer a framework in which each distribution parameter can be modeled separately by its own GAM. In Figure~\ref{Devel_parmod} the development of parametric models starting from simple linear models (LM) and ranging to GAMLSS is illustrated.\\
These flexible parametric methods provide well fitting models when applied on precipitation data in a rather smooth/homogeneous setting, e.g. areas where no abrupt changes in meteorological/climatological events appear. However, for example in complex terrain/mountainous regions various factors can change rapidly and their effects are difficult to be modeled/represented by a smooth parametric model. In this situation a different approach which can deal with these abrupt changes would be to apply tree structured models such as regression trees (\citealt{loh2011classification}) or random forests (\citealt{breiman2001random}). Moreover, these models are particularly advantageous in case of many covariates with unknown effects or interactions as the variable selection is done automatically within the model fitting process. Therefore, no additional effort or knowledge for the specification of the model is required in advance. However, even though their main advantage of being able to model non-additive effects and interactions without any necessary prespecification can play to their strength, regression trees are not an adequate choice if (additional) smooth effects appear as well. In this situation a random forest provides a good compromise as it combines a set of trees such that the steps imposed by the trees are smoothed out in case of smooth effects while abrupt changes are still captured by the tree structure of the model.\\
For these reasons, regression trees and particularly random forests provide valuable features which can contribute to improving models for probabilistic forecasting. But, up to now, no general distributional framework is available for these methods. Therefore, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. The developments in the field of tree-structured models together with the novel approaches of distributional trees and forests are illustrated in Figure~\ref{Devel_treeforest}.\\

@@

<<plot_motivation_disttree, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
lines(x = x, y = ytree, col = pal["forest"], lwd=7, main = "")
polygon(c(x, rev(x)), c(ytree + var, rev(ytree - var)),
  col = pallight["forest"], border = "transparent")
box(lwd=5)
@
\vspace{-0.2cm}%\\
Distributional Tree
\vspace{0.2cm}
\end{center}
\endminipage};
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](2.1,0)--(2.9,0);
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](7.1,0)--(7.9,0);
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](0,-2.4)--(0,-3.2);
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](2.2,-5.4)--(7.9,-2.5);
\end{tikzpicture}
\caption{\label{Devel_treeforest}Development in the field of algorithmic trees and forests}
\end{figure}

@@

@@

In that way, a whole distribution is specified with all its features, including location, scale and shape. However, fitting one global model to the whole data set might be too much of a generalization and therefore not represent the data and its features in a reasonable way. Moreover, if covariates are available it is desirable to include them into the model as they might provide important additional information. In our application on precipitation data a wide range of covariates are available from numerical ensemble predictions considering precipitation, temperature, sea level pressure and other meteorological quantities which might be of interest/have an influence on the predictions. \\

@@

\subsection{Distributional Tree}
\label{disttree}
Given the fit of a global distributional model $D(Y, \bold{\hat{\theta}})$ the idea is to use the information provided by this model together with a set of covariates $\bold{Z} = (Z_1, \ldots, Z_m) \in \mathcal{Z}$ (from now on denoted by $\bold{Z}$ instead of $\bold{X}$ to indicate that they are considered as split variables only and not as regressors) to decide whether the data set should be split into subgroups or not. In particular, the information provided by the fitted model are the scores 
$$
s(\hat{\theta}, y_i) = \frac{\partial \ell}{\partial \theta}(\hat{\theta}; y_i).
$$ 
These values express how well the model fits the data regarding each parameter. Ideally they should fluctuate randomly around zero, similar to residuals in ordinary least squares estimations, but in this case there is a score value for each pair of estimated parameter $\hat{\theta}_j \in (\hat{\theta}_1, \ldots, \hat{\theta}_k) = \bold{\hat{\theta}}$ and observation $y_j \in \{y_i\}_{i = 1,\ldots, n}$. This enables the model to be sensitive to changes in each of the parameters which is one of the main advantages of distributional trees over other tree models considering only the mean of the distribution.\\
Statistical tests are applied to check whether there is a significant connection/dependency between the scores and each of the partitioning variables $Z_l \in \bold{Z} = (Z_1, \ldots, Z_m)$. If so, the data set is split into subgroups based on the values of the partitioning variable that shows the highest connection/dependency.\\
In a next step a local distributional model is fit in each of the subgroups. Repeating this procedure of fitting a model, applying statistical tests to evaluate the goodness of fit and splitting the data set depending on the test results leads to a tree structure. In that way the learning data set is separated into \textit{B} disjoint segments such that
$$
\mathcal{Z} = \dot{\bigcup\limits_{b = 1, \ldots, \textit{B}}} \mathcal{B}_b
$$
where each segment is represented by a terminal node of the tree.\\

The exact way of determining how and where to split varies across different tree algorithms. For distributional trees, one of two algorithms can be chosen as a framework to build the tree: either the MOB algorithm (\citealt{zeileis2008model}) or the ctree algorithm (\\citealt{hothorn2006unbiased}). Summed up, the steps of building a distributional tree using one of these algorithms are
\begin{enumerate}
\item Specify a distribution with log-likelihood function $\ell(\theta; y)$.
\item Estimate $\bold{\hat{\theta}}$ via maximum likelihood.
\item Test for associations or instabilities of the scores $s(\bold{\hat{\theta}}, y_i)$ and each partitioning variable $Z_l$.
%\item Assess whether the \emph{model scores} are associated with
%    (or change along) any of the available covariates -- e.g.,
%    using parameter instability tests (\emph{strucchange}) or
%    conditional inference (\emph{coin}).
\item Split the sample along the partitioning variable with the strongest association or instability.
    Choose the breakpoint with the highest improvement in log-likelihood.
\item Repeat steps 2--4 recursively in the subgroups
    until some stopping criterion is reached.
    %-- e.g., for significance or sample size.
%\item Choose the variable with the strongest association
%\item Choose the split point which leads to the highest improvement of the model
%\item Split and repeat 2-6 in each node until a stopping criterion is met
\end{enumerate}
The stopping criterion in step 5 can for example be a minimum number of observations in a node or a p-value for the statistical tests applied in step 3.\\
Executing these steps results in a stepwise parametric model where a complete distribution is fit in each node of the tree.\\
\\
For distributional trees the strategy of how to make a prediction for a new set of covariates $\bold{z}^{\ast} = (z_1^{\ast}, \ldots, z_m^{\ast})$ is very simple and straight forward. All observations of the learning data set $\{(y_i,\bold{z}_i)\}_{i=1\ldots,n}$ which end up in the same terminal node of the tree as the new set of covariates are selected and a distributional model is fit on the subset. To select this subset a weight vector $\bold{\omega}^{tree} = (\omega^{tree}_1, \ldots, \omega^{tree}_n)$ is included in the fitting process.
$$
\omega^{tree}_i(\bold{z}^{\ast}) = \sum_{b=1}^B I((\bold{z}_i \in \mathcal{B}_b) \land (\bold{z}^{\ast} \in \mathcal{B}_b))
$$
The predicted distribution of the new set of covariates is then fully specified by the estimated parameter $\hat{\theta}(\bold{z}^{\ast})$ where
$$
\hat{\theta}(\bold{z}^{\ast}) = \max_{\theta \in \Theta} \sum_{i=1}^n \omega^{tree}_i(\bold{z}^{\ast}) \cdot \ell(\theta; y_i).
$$

While the clear structure of such a tree model is advantageous in interpreting and analyzing it, the abrupt changes are often too rough and impose steps on the model even if the true underlying effect is smooth. In these situations bagging can solve the problem by combining an ensemble of trees such that wrongly detected abrupt changes of the model are turned into almost smooth transitions. In that way, all kinds of effects can be modeled, which is is also one the of the main advantages of forest models (taking bagging one step further as the variance is reduced by sampling the covariates considered for the splitting process in each node). \\
\\
Common forest models such as random forests consist of trees only considering the mean of the distribution of the response variable. For that reason, changes in other parameters can only be detected if they are connected to changes in the location parameter, e.g. if they depend on the location parameter. Therefore, building a forest of distributional trees which can also capture separate changes in each of the distribution parameters leads to an improvement of the model. This valuable feature of detecting separate changes is also provided by GAMLSS as each parameter is modeled separately. However, in this case this separation can also be a disadvantage as it hinders the capturing of interactions unless they are specified explicitly by the user. These two examples/situations provide once again a good reason to combine the two frameworks of flexible parametric modeling and tree structured models in order to profit from their advantages and compensate possible disadvantages. This is the pursued goal in introducing distributional forests.

@@

\subsection{Models and Evaluation}
Next to distributional forests three other methods are applied to this data set and their results are compared based on CRPS values. One of the reference models is an EMOS models as this is a common choice in this type of setting / research field. The other two models are both GAMs, once in a prespecified form in terms of variable selection for each distribution parameters and once applying a boosting algorithm). 
The chosen methods and their specifications are listed in Table~\ref{model_specification}.
\begin{table}[t!]
%\resizebox{1.4\textwidth}{!}{
%\begin{minipage}{\textwidth}
\begin{center}
\begin{tabular}{ l  l   l   l }
\hline
Model & Type & Location ($\mu$) & Scale ($\log(\sigma)$) \\
\hline
\textbf{Forest} & recursive    & all & all \\ 
                & partitioning &     &     \\ 
\hline
\textbf{prespecified GAM} & spline  & tp{\_}mean, & tppow{\_}sprd,\\
                            & in each & tp{\_}max,  & dswrf{\_}sprd{\_}mean,\\
 & & tp{\_}mean1218 $\ast$ & tp{\_}sprd1218 $\ast$\\ 
 & & \quad cape{\_}mean1218, & \quad  cape{\_}mean1218,\\
 & & dswrf{\_}mean{\_}mean, & tcolc{\_}sprd{\_}mean,\\
 & & tcolc{\_}mean{\_}mean, & tdiff500850{\_}mean\\
 & & pwat{\_}mean{\_}mean, & \\
 & & tdiff500850{\_}mean, & \\
 & & msl{\_}diff & \\
\hline
\textbf{boosted GAM} & spline  & all & all \\
                       & in each &     &     \\
\hline
\textbf{EMOS} & linear & tp{\_}mean & tp{\_}spread \\
\hline
\end{tabular}
\caption[Table caption text]{Table of applied methods together with their type (way of including the covariates) and the included covariates for each distribution parameter.}
\label{model_specification}
%}
\end{center}
\end{table}

@@

par(mfrow = c(2,2))
pithist(pit_df_t, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "Forest")
pithist(pit_g_t, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "prespecified GAM")
pithist(pit_gb_t, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "boosted GAM")
pithist(pit_ml_t, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "EMOS")

@@

The CRPS skills score boxplot shows that, in this setting, both GAM models and the forest perform better than the EMOS model. While the two GAM models lead to an improvement of around 4 percent with the boosted version being slightly ahead, the forest even reaches an improvement rate of close to 6 percent.\\
From this result we can conclude that the distributional forest provides an alternative framework where only little or almost no effort is required to set up the model and which is competitive to existing methods that are commonly used for this kind of application or even outperforms them.\\
For the specified GAM model the covariates needed to be selected and interactions needed to be defined in advance. To do so, meteorological experience and knowledge is required and a lot of thought has to be put in these decisions. Moreover, making such decisions is always equivalent to influencing the resulting model and comes along with the risk of having missed important facts or even imposing certain effects on the model which might not match / correspond to the truth.\\
This risk is avoided in the boosted GAM as all covariates are included and the selection of the influential covariates is done within a boosting framework. However, the trade off here is a remarkably higher computational effort.\\
To build a distributional forest the user does not have to put any additional effort in it and therefore is not required to provide any expert knowledge or additional information about the setting. Variable selection is done automatically and interactions are detected by the tree structure of the model.\\
In this particular setting, a very high number of covariates is available which intensifies / increases the impact of the above mentioned strength of the forest model.\\
Therefore, in comparing these methods one advantage of distributional forests is clearly outlined as they offer a very flexible framework which enables an easy use of the method and at the same time lead to an improvement in the predictive performance.

@@

Figure~\ref{map} shows the location of all stations in the data set. Each observation station is represented by a point colored according to the method that performed best in terms of CRPS at this station. For those points where the distributional forest is not the best performing method the intensity of the color represents how much better (in terms of CRPS) the method is compared to the distributional forest. The more intense the color, the bigger the difference. In case of the forest being the best model, the second best method is chosen as reference.

@@

\section*{Computational Details}
The corresponding implementation to the novel methods presented in this article is available on R-Forge in an \textsf{R} package called \textbf{disttree} within the project \textbf{partykit} (\url{https://R-Forge.R-project.org/projects/partykit/}). The function \code{disttree} uses either the \code{mob} or \code{ctree} function as a framework to build a tree. Similarly the function \code{distforest} is based on the \code{cforest} function which provides a framework to build a forest. All three functions, \code{mob}, \code{ctree} and \code{cforest}, are provided in the package \textbf{partykit} (\citealt{hothorn2015partykit}).

>>>>

\Abstract{
In the field of probabilistic forecasting linear models such as ensemble model output statistics (EMOS, \citealt{gneiting2005calibrated}) provide common and widely used tools. This is due to their simplicity and interpretability. However, the restriction to model only linear effects requires further extensions which allow for more flexibility. Generalized additive models for location, scale, and shape (GAMLSS, \citealt{stasinopoulos2005}) can also capture non-linear effects as each distribution parameter can be modeled separately by its own generalized additive model (GAM). A different approach that would make it possible to also model non-additive effects and interactions is to apply tree structured models such as regression trees (\citealt{loh2011classification}) or random forests (\citealt{breiman2001random}). However, up to now, this class of models has not yet been integrated in probabilistic forecasting as no general distributional framework is available for these methods. Therefore, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. Applying these novel approaches on precipitation data in a mountainous region with a large number of numerical weather prediction quantities illustrates their strengths as their performance is compared to GAMLSS models that are specified based either on subject-matter knowledge and experience from meteorology or based on a computationally more demanding boosting approach.

@@

\SweaveOpts{eval = TRUE}
\section{Introduction}
A typical application in probabilistic forecasting in meteorology is the analysis of precipitation data. The main interest lies in estimating the probability of precipitation as well as predicting the expected amount of precipitation. In this research area various methods have been developed and tested in the last decades. (Multiple) Linear models such as ensemble model output statistics (EMOS) were introduced more than a decade ago (\citealt{gneiting2005calibrated}) and are still common and widely used. Despite their simplicity and interpretability, the limitation to modeling only linear effects has risen the interest in applying more flexible models. This feature is provided by generalized additive models (GAM) which can also model non-linear effects. However, often only the expectation of the underlying distribution is estimated. To obtain probabilistic predictions, generalized additive models for location, scale, and shape (GAMLSS, \citealt{stasinopoulos2005}) offer a framework in which each distribution parameter can be modeled separately by its own GAM. In Figure~\ref{Devel_parmod} the development of parametric models starting from simple linear models (LM) and ranging to GAMLSS is illustrated.\\
These flexible parametric methods provide well fitting models when applied on precipitation data in a rather smooth and homogeneous setting, e.g. areas where no abrupt changes in meteorological/climatological events appear. However, for example in complex terrain in mountainous regions various factors can change rapidly and their effects are difficult to be represented by a smooth parametric model. In this situation a different approach which can deal with these abrupt changes would be to apply tree structured models such as regression trees (\citealt{loh2011classification}) or random forests (\citealt{breiman2001random}). Moreover, these models are particularly advantageous in case of many covariates with unknown effects or interactions as the variable selection is done automatically within the model fitting process. Therefore, no additional effort or knowledge for the specification of the model is required in advance. However, even though their main advantage of being able to model non-additive effects and interactions without any necessary prespecification can play to their strength, regression trees are not an adequate choice if (additional) smooth effects appear as well. In this situation a random forest provides a good compromise as it combines a set of trees such that the steps imposed by the trees are smoothed out in case of smooth effects while abrupt changes are still captured by the tree structure of the model.\\
For these reasons, regression trees and particularly random forests provide valuable features which can contribute to improving models for probabilistic forecasting. But, up to now, no general distributional framework is available for these methods. Therefore, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. The developments in the field of tree-structured models together with the novel approaches of distributional trees and forests are illustrated in Figure~\ref{Devel_treeforest}.\\

@@

<<plot_motivation_disttree, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
lines(x = x, y = ytree, col = pal["forest"], lwd=7, main = "")
polygon(c(x, rev(x)), c(ytree + var, rev(ytree - var)),
  col = pallight["forest"], border = "transparent")
box(lwd=5)
@
\vspace{-0.2cm}%\\
Distributional Tree
\vspace{0.2cm}
\end{center}
\endminipage};
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](2.1,0)--(2.9,0);
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](7.1,0)--(7.9,0);
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](0,-2.4)--(0,-3.2);
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](2.2,-5.4)--(7.9,-2.5);
\end{tikzpicture}
\caption{\label{Devel_treeforest}Development in the field of algorithmic trees and forests.}
\end{figure}

@@

\newpage

@@

In that way, a whole distribution is specified with all its features, including location, scale and shape. However, fitting one global model to the whole data set might be too much of a generalization and therefore not represent the data and its features in a reasonable way. Moreover, if covariates are available it is desirable to include them into the model as they might provide important additional information. In our application on precipitation data a wide range of covariates are available from numerical ensemble predictions considering precipitation, temperature, sea level pressure and other meteorological quantities which might be of interest and have an influence on the predictions. \\

@@

\subsection{Distributional Tree}
\label{disttree}
Given the fit of a global distributional model $D(Y, \bold{\hat{\theta}})$ the idea is to use the information provided by this model together with a set of covariates $\bold{Z} = (Z_1, \ldots, Z_m) \in \mathcal{Z}$ (from now on denoted by $\bold{Z}$ instead of $\bold{X}$ to indicate that they are considered as split variables only and not as regressors) to decide whether the data set should be split into subgroups or not. In particular, the information provided by the fitted model are the scores 
$$
s(\hat{\theta}, y_i) = \frac{\partial \ell}{\partial \theta}(\hat{\theta}; y_i).
$$ 
These values express how well the model fits the data regarding each parameter. Ideally they should fluctuate randomly around zero, similar to residuals in ordinary least squares estimations, but in this case there is a score value for each pair of estimated parameter $\hat{\theta}_j \in (\hat{\theta}_1, \ldots, \hat{\theta}_k) = \bold{\hat{\theta}}$ and observation $y_j \in \{y_i\}_{i = 1,\ldots, n}$. This enables the model to be sensitive to changes in each of the parameters which is one of the main advantages of distributional trees over other tree models considering only the mean of the distribution.\\
Statistical tests are applied to check whether there is a significant dependency between the scores and each of the partitioning variables $Z_l \in \bold{Z} = (Z_1, \ldots, Z_m)$. If so, the data set is split into subgroups based on the values of the partitioning variable that shows the highest dependency.\\
In a next step a local distributional model is fit in each of the subgroups. Repeating this procedure of fitting a model, applying statistical tests to evaluate the goodness of fit and splitting the data set depending on the test results leads to a tree structure. In that way the learning data set is separated into \textit{B} disjoint segments such that
$$
\mathcal{Z} = \dot{\bigcup\limits_{b = 1, \ldots, \textit{B}}} \mathcal{B}_b
$$
where each segment is represented by a terminal node of the tree.\\

The exact way of determining how and where to split varies across different tree algorithms. For distributional trees, one of two algorithms can be chosen as a framework to build the tree: either the MOB algorithm (\citealt{zeileis2008model}) or the ctree algorithm (\citealt{hothorn2006unbiased}). Summed up, the steps of building a distributional tree using one of these algorithms are
\begin{enumerate}
\item Specify a distribution with log-likelihood function $\ell(\theta; y)$.
\item Estimate $\bold{\hat{\theta}}$ via maximum likelihood.
\item Test for associations or instabilities of the scores $s(\bold{\hat{\theta}}, y_i)$ and each partitioning variable $Z_l$.
%\item Assess whether the \emph{model scores} are associated with
%    (or change along) any of the available covariates -- e.g.,
%    using parameter instability tests (\emph{strucchange}) or
%    conditional inference (\emph{coin}).
\item Split the sample along the partitioning variable with the strongest association or instability.
    Choose the breakpoint with the highest improvement in log-likelihood.
\item Repeat steps 2--4 recursively in the subgroups
    until some stopping criterion is reached.
    %-- e.g., for significance or sample size.
%\item Choose the variable with the strongest association
%\item Choose the split point which leads to the highest improvement of the model
%\item Split and repeat 2-6 in each node until a stopping criterion is met
\end{enumerate}
The stopping criterion in step 5 can for example be a minimum number of observations in a node or a p-value for the statistical tests applied in step 3.\\
Executing these steps results in a stepwise parametric model where a complete distribution is fit in each node of the tree.\\
\\
For distributional trees the strategy of how to make a prediction for a new set of covariates $\bold{z}^{\ast} = (z_1^{\ast}, \ldots, z_m^{\ast})$ is very simple and straight forward. All observations of the learning data set $\{(y_i,\bold{z}_i)\}_{i=1\ldots,n}$ which end up in the same terminal node of the tree as the new set of covariates are selected and a distributional model is fit on the subset. To select this subset a weight vector $\bold{\omega}^{tree} = (\omega^{tree}_1, \ldots, \omega^{tree}_n)$ is included in the fitting process.
$$
\omega^{tree}_i(\bold{z}^{\ast}) = \sum_{b=1}^B I((\bold{z}_i \in \mathcal{B}_b) \land (\bold{z}^{\ast} \in \mathcal{B}_b))
$$
The predicted distribution of the new set of covariates is then fully specified by the estimated parameter $\hat{\theta}(\bold{z}^{\ast})$ where
$$
\hat{\theta}(\bold{z}^{\ast}) = \max_{\theta \in \Theta} \sum_{i=1}^n \omega^{tree}_i(\bold{z}^{\ast}) \cdot \ell(\theta; y_i).
$$

While the clear structure of such a tree model is advantageous in interpreting and analyzing it, the abrupt changes are often too rough and impose steps on the model even if the true underlying effect is smooth. In these situations bagging can solve the problem by combining an ensemble of trees such that wrongly detected abrupt changes of the model are turned into almost smooth transitions. In that way, all kinds of effects can be modeled, which is is also one the of the main advantages of forest models (taking bagging one step further as the variance is reduced by sampling the covariates considered for the splitting process in each node). \\
\\
Common forest models such as random forests consist of trees only considering the mean of the distribution of the response variable. For that reason, changes in other parameters can only be detected if they are connected to changes in the location parameter, e.g. if they depend on the location parameter. Therefore, building a forest of distributional trees which can also capture separate changes in each of the distribution parameters leads to an improvement of the model. This valuable feature of detecting separate changes is also provided by GAMLSS as each parameter is modeled separately. However, in this case this separation can also be a disadvantage as it hinders the capturing of interactions unless they are specified explicitly by the user. These two cases provide once again a good reason to combine the two frameworks of flexible parametric modeling and tree structured models in order to profit from their advantages and compensate possible disadvantages. This is the pursued goal in introducing distributional forests.

@@

\subsection{Models and Evaluation}
Next to distributional forests three other methods are applied to this data set and their results are compared based on CRPS values. One of the reference models is an EMOS models as this is a common choice in this type of setting and research field. The other two models are both GAMs, once in a prespecified form in terms of variable selection for each distribution parameters and once applying a boosting algorithm). 
The chosen methods and their specifications are listed in Table~\ref{model_specification}.
\begin{table}[t!]
%\resizebox{1.4\textwidth}{!}{
%\begin{minipage}{\textwidth}
\begin{center}
\begin{tabular}{ l  l   l   l }
\hline
Model & Type & Location ($\mu$) & Scale ($\log(\sigma)$) \\
\hline
\textbf{Forest} & recursive    & all & all \\ 
                & partitioning &     &     \\ 
\hline
\textbf{prespecified GAM} & spline  & tp{\_}mean, & tppow{\_}sprd,\\
                            & in each & tp{\_}max,  & dswrf{\_}sprd{\_}mean,\\
 & & tp{\_}mean1218 $\ast$ & tp{\_}sprd1218 $\ast$\\ 
 & & \quad cape{\_}mean1218, & \quad  cape{\_}mean1218,\\
 & & dswrf{\_}mean{\_}mean, & tcolc{\_}sprd{\_}mean,\\
 & & tcolc{\_}mean{\_}mean, & tdiff500850{\_}mean\\
 & & pwat{\_}mean{\_}mean, & \\
 & & tdiff500850{\_}mean, & \\
 & & msl{\_}diff & \\
\hline
\textbf{boosted GAM} & spline  & all & all \\
                       & in each &     &     \\
\hline
\textbf{EMOS} & linear & tp{\_}mean & tp{\_}spread \\
\hline
\end{tabular}
\caption[Table caption text]{Table of applied methods together with their type (way of including the covariates) and the included covariates for each distribution parameter.}
\label{model_specification}
%}
\end{center}
\end{table}

@@

par(mfrow = c(2,2))
pithist(pit_df_t, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "Forest", ylim = c(0,1.4))
pithist(pit_g_t, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "prespecified GAM", ylim = c(0,1.4))
pithist(pit_gb_t, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "boosted GAM", ylim = c(0,1.4))
pithist(pit_ml_t, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "EMOS", ylim = c(0,1.4))

@@

The CRPS skills score boxplot shows that, in this setting, both GAM models and the forest perform better than the EMOS model. While the two GAM models lead to an improvement of around 4 percent with the boosted version being slightly ahead, the forest even reaches an improvement rate of close to 6 percent.\\
From this result we can conclude that the distributional forest provides an alternative framework where only little or almost no effort is required to set up the model and which is competitive to existing methods that are commonly used for this kind of application or even outperforms them.\\
For the specified GAM model the covariates needed to be selected and interactions needed to be defined in advance. To do so, meteorological experience and knowledge is required and a lot of thought has to be put in these decisions. Moreover, making such decisions is always equivalent to influencing the resulting model and comes along with the risk of having missed important facts or even imposing certain effects on the model which might not correspond to the truth.\\
This risk is avoided in the boosted GAM as all covariates are included and the selection of the influential covariates is done within a boosting framework. However, the trade off here is a remarkably higher computational effort.\\
To build a distributional forest the user does not have to put any additional effort in it and therefore is not required to provide any expert knowledge or additional information about the setting. Variable selection is done automatically and interactions are detected by the tree structure of the model.\\
In this particular setting, a very high number of covariates is available which intensifies the impact of the above mentioned strength of the forest model.\\
Therefore, in comparing these methods one advantage of distributional forests is clearly outlined as they offer a very flexible framework which enables an easy use of the method and at the same time lead to an improvement in the predictive performance.

@@

Figure~\ref{map} shows the location of all stations in the data set. Each observation station is illustrated by a symbol with the type of symbol representing the method that performed best in terms of CRPS at this station. The color of the symbol indicates the difference of the CRPS of the forest model and the CRPS of the best perfoming among the other three models (prespecified GAM, boosted GAM and EMOS).

@@

\section*{Computational Details}
The corresponding implementation to the novel methods presented in this article is available on R-Forge in an \textsf{R} package called \textbf{disttree} within the project \textbf{partykit} (\url{https://R-Forge.R-project.org/projects/partykit/}). The function \code{disttree} uses either the \code{mob} or \code{ctree} function as a framework to build a tree. Similarly, the function \code{distforest} is based on the \code{cforest} function which provides a framework to build a forest. All three functions, \code{mob}, \code{ctree} and \code{cforest}, are provided in the package \textbf{partykit} (\citealt{hothorn2015partykit}).
